---
title: "Transform Data"
---

## Load data

```{r Your Turn 1, message=FALSE}
library(tidyverse)  # Provides functions used throughout this session

covid_testing <- read_csv("data/covid_testing.csv") #Read in data
```

## Convert steps into a pipeline with the pipe

```{r Steps and the pipe}
# These 3 lines of code are an example of the how we would accomplish a multi-step analysis based on what we've learned so far. 
day_10 <- filter(covid_testing, pan_day <= 10)
day_10 <- select(day_10, clinic_name)
day_10 <- arrange(day_10, clinic_name)
day_10

# vs

# This code accomplishes the same task and is far more legible. 
covid_testing %>% 
  filter(pan_day <= 10) %>% 
  select(clinic_name) %>% 
  arrange(clinic_name)
```

## Your Turn 1

Use `%>%` in the following chunk to write a sequence of functions that: 

1. Filters to orders from the clinic (`clinic_name`) of "picu"

2. Selects the columns with the receive to verify turnaround time (`rec_ver_tat`) and day from start of the pandemic (`pan_day`)

3. Arrange the `pan_day` from highest to lowest so you can see most recent tests on top.

4. Use <- to assign the result to a new variable and take a look.

CHALLENGE: Use your mouse to select the name of the new data frame from the list of objects in the Environment tab in the upper-right pane of Rstudio. Answer the following questions:

- Can you find the order whose TAT exceeded 10 hours?
- What is the typical TAT for the PICU?
- Given these data does it appear that there is a chronic problem in TAT for the PICU, or was this test an outlier?

```{r Your Turn 1, eval = FALSE}
covid_testing %>% 
  filter(________) %>% # filter to the clinic with the name "picu"
  select(________) %>% # select the columns you want to keep
  arrange(________)    # arrange the pan_day column from highest to lowest (check slide 42 the arrange section of this file to use the desc() function)
```

## mutate

```{r}
# The following code creates a data frame with a new column called c_r_tat_mins which contains the collect to receive TAT quantified in minutes.
covid_testing %>%
	mutate(c_r_tat_mins = col_rec_tat * 60)
```

## Your Turn 2
Create a new column using the mutate() function that contains the total TAT. 
Hint: The total TAT is the sum of `col_rec_tat` and `rec_ver_tat`

```{r Your Turn 2, eval=FALSE}
covid_testing %>%
	mutate(col_ver_tat = _________)
```

## Replace a column with `mutate()`

```{r}
covid_testing %>%
	mutate(mrn = as.character(mrn))
```


# Takeaways

* Connect operations with `%>%`  
* Make new variables, with `mutate()`  
